---
title: "Location geocoding workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Location geocoding workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE, purl = FALSE)
```

`locatr` turns a messy address table into an audit-ready crosswalk. The chunks
below are not evaluated because they need live geocoding services, but they show
the intended end-to-end run. The package assumes you already have a data frame;
it does not connect to databases or source systems.

## 1. Pull and clean

```{r}
library(locatr)

cleaned <- records %>%
  clean_addresses(
    id = `Location ID`, address = Address,
    city = City, zip = Zip, name = `Location Name`
  ) %>%
  flag_bad_addresses()
```

`clean_addresses()` adds `*_clean` columns and a `full_address_clean`
string; `flag_bad_addresses()` sends PO boxes and placeholders straight to
review. Only address and city are required. If your file has no ID, locatr
generates row-number IDs; if it has no ZIP, `zip_clean` stays `NA` and the
single-line address omits the ZIP instead of ending in `NA`:

```{r}
cleaned_minimal <- records %>%
  clean_addresses(address = Address, city = City, state = "NJ") %>%
  flag_bad_addresses()
```

Missing ZIP is recorded as `bad_address_flag == "missing_zip"` for audit, but it
does not block geocoding when address + city + state are present.

For one-off review, use `geocode_address()` to see ranked ArcGIS candidates for
a single address:

```{r}
if (interactive()) {
  geocode_address("1600 Pennsylvania Ave NW", city = "Washington", state = "DC")
}
```

## 2. Geocode with a guarded cascade

```{r}
geocoded <- geocode_records(cleaned)   # Census -> ArcGIS -> name lookup
```

`geocode_records()` runs each tier in turn and validates against the
configured region after each pass, so a later, fuzzier service only retries what
is still unplaced. The `geocode_pass` column records which tier placed each row.

## 3. Geography join

```{r}
with_geography <- add_county_muni(geocoded, state = "NJ")
```

The geography step is independent of the input address columns once coordinates
exist. `add_county_muni()` builds Census TIGER/Line geography and attaches
county/locality fields. Pass an `sf` boundary layer to adapt the geography join,
e.g. `add_muni_from_shapes(geocoded, muni_shapes = my_local_shapes)`, or use
`add_muni_from_key()` when your records and geography share a code column.

## 4. Review, override, export

```{r}
write_geocode_review(with_geography, "manual_review.csv")

final <- with_geography %>%
  apply_manual_overrides("manual_review_completed.csv") %>%
  export_location_crosswalk("location_crosswalk.csv")
```

`final` is ready for Tableau, GIS joins, or a reusable reference table, and
every row carries the audit columns that explain how its coordinate was
produced.
